 

### 2021-06-07 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

While a few studies show $Q_{10}$ to be greater in upper-canopy leaves than in shaded lower-canopy leaves across species [@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996]. Contranstingly, in a lowland tropical Australian forest study, $Q_{10}$ did not vary across species vertical profiles [@weerasingheCanopyPositionAffects2014], similary across temperate trees at Coweeta, though with a slight tendency to increase down the canopy relative to upper-canopy leaves as $T_{air}$ increased (*see their Fig. 1*)[@bolstad_foliar_1999]. (*Martijn has some unpublished data on this that we might include.unpublished data showing higher Q10 in shade than in sun leaves for 10 species or so. May make sense to include here: https://github.com/EcoClimLab/vertical-thermal-review/issues/8.*). @bolstad_foliar_1999 also found evidence of acclimation, but there remained a declining trend of respiration with elevation.
- Response of respiration to $T_{leaf}$ [partially represented  by $E_{0}$] in Quercus rubra hardly varies with canopy position, but canopy-position effect in respiration is much greater in mesic sites, and seasonal effect is greater in drier sites [@xuSeasonalVariationTemperature2006]

@@

Canopies are crucial to microclimatic functioning, as they buffer several aspects of the microenvironment. This buffering is influenced by the hydraulogical-cycle at large, therefore severe droughts place canopies at risk, particularly larger trees that suffer more during droughts. Further, increase in global temperatures has shown that canopies of temperate and tropical forests function close to thermal threshold and beyond photosynthetic optimas [@mau_temperate_2018; @huangAirTemperatureOptima2019]. Increase in such canopy $T_{leaf}$ reduces forest GPP, specifically in the tropics where thermal limits are narrower than temperate forests.

@@

While climate change driven forest shifts are observed, certain forests are more resilient to warming and droughts than others, such as African rainforests [Amy Bennett's paper]. Forest resiliency is also observed through increasing temperature acclimation [martijn's paper and others] and reversible recovery of canopy heat stress [Smith et al. 2020] among other mechanisms [for eg. *REF*]

>>>>

While a few studies show $Q_{10}$ to be greater in upper-canopy leaves than in shaded lower-canopy leaves across species [@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996]. Contrastingly, in a lowland tropical Australian forest study, $Q_{10}$ did not vary across species vertical profiles [@weerasingheCanopyPositionAffects2014], similary across temperate trees at Coweeta though with a slight tendency to increase down the canopy relative to upper-canopy leaves as $T_{air}$ increased (*see their Fig. 1*)[@bolstad_foliar_1999]. (*Martijn has some unpublished data on this that we might include.unpublished data showing higher Q10 in shade than in sun leaves for 10 species or so. May make sense to include here: https://github.com/EcoClimLab/vertical-thermal-review/issues/8.*). @bolstad_foliar_1999 also found evidence of acclimation, but there remained a declining trend of respiration with elevation.

Further, @xuSeasonalVariationTemperature2006 showed that the response of respiration to $T_{leaf}$ [partially represented  by $E_{0}$] in Quercus rubra hardly varied with canopy position. However, considering site differences, canopy-position effect in respiration was much greater in mesic sites, and seasonal effect is greater in drier sites.

@@

Canopies are crucial to microclimatic functioning, as they buffer several aspects of the microenvironment [Zelleweger et al. 2019]. This buffering is influenced by the hydraulogical-cycle at large [Davis et al. 2019], therefore severe droughts place canopies at risk, particularly larger trees that suffer more during droughts[Bennett et al. 2015]. Further, increase in global temperatures has shown that canopies of temperate and tropical forests function close to thermal threshold and beyond photosynthetic optimas [@mau_temperate_2018; @huangAirTemperatureOptima2019]. Increase in such canopy $T_{leaf}$ reduces forest GPP [Pau et al. 2018], specifically in the tropics where thermal limits are narrower than temperate forests [Perez and Feeley, 2018].[**to include Teskey 2015; Breashers et al. 2021**]

@@

While climate change driven forest shifts are observed, certain forests are more resilient to warming and droughts than others, such as African rainforests [Amy Bennett's new paper]. Forest resiliency is also observed through increasing temperature acclimation among tropical and boreal species [martijn's papers; Danielle Way's papers; Atkin's papers] and reversible recovery of canopy heat stress [Smith et al. 2020] among other mechanisms [for eg. *REF*]
